Exception Handling Bug Hazards in Android

### 2015-10-07 SPRINGER_JOURNAL/msr_journal.tex
survey section changes

\title{Exception Handling Bug Hazards in Android}
\subtitle{Results from a Mining and Survey-based Study}

@@

\institute{F. Author \at
              first address \\
              Tel.: +123-45-678910\\
              Fax: +123-45-678910\\
              \email{fauthor@example.com}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           S. Author \at
              second address
}

@@

\begin{abstract}
Insert your abstract here. Include keywords, PACS and mathematical
subject classification numbers as needed.
\keywords{First keyword \and Second keyword \and More}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

In recent years, we have witnessed an astonishing increase in the number of
mobile applications. These applications extend phones capabilities 
far beyond basic calls and textual messages. They must, however,
face an increasing number of threats to application robustness
 arising from failures in the underlying middleware and hardware (e.g., camera, sensors);
failures in third party services and libraries; compatibility issues~\cite{McDon13}; 
 memory and batery restrictions; and noisy external resources~\cite{Zhang12} 
(e.g., wireless connections,GPS, bluetooth).

Therefore, techniques for error detection and handling are not  an optional add-on but a 
fundamental part of such apps. The exception handling mechanism~\cite{goodenough1975exception},
embedded in many mainstream programming languages, such as Java, C++ and C\#,
 is one of the most used techniques for detecting and recovering from such exceptional conditions.
In this paper we will be concerned with exception handling in Android apps,
which reuses Java's exception handling model.

@@

Studies have shown that exception-related code is generally poorly understood and among the least tested parts of the system~\cite{miller1997issues,Robil00,shah2010understanding, 
garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling,yuan:2014.osdi}.
As a consequence they may inadvertently negatively affect the system: exception-related code may introduce failures such as 
uncaught exceptions~\cite{jo2004uncaught, Zhang12} - 
which can lead to system crashes, making the system even less robust~\cite{coelho2011unveiling}.

In Java, when an application fails due to an uncaught exception, 
it automatically terminates, while the system prints a stack trace to the console, 
or on a log file~\cite{gosling2000java}.  A typical Java stack trace consists of  the fully qualified name 
of the thrown exception and the ordered list of methods that were active on the call stack before 
the exception occurred~\cite{gosling2000java,bloch2008effective}.

@@

%When available, the exception stack traces provide a useful source of information about system crashes ~\cite{bettenburg2008makes} which 
% can enable different kinds of post-mortem analysis and support:  debugging~\cite{schroter2010stack}, 
% bug classification and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},  
%automated bug fixing~\cite{sinha2009fault} and fault-proneness prediction models~\cite{kim2013predicting}.

This study performs a post mortem analysis of the exception stack traces 
included in issues
reported on Android projects hosted in GitHub and Google Code. 
The goal of this study is to investigate whether the reported exception stack traces
 can reveal common \emph{bug hazards} in the exception-related code.  
A \emph{bug hazard}~\cite{binder2000testing}  is a circumstance that increases the 
chance of a bug being present in the software. An example of a bug hazard can 
be a characteristic of  the exception-related code which can increase the likelihood 
of introducing the aforementioned uncaught exceptions.

@@

 The high prevalence of NullPointerExceptions is in line with findings of earlier research~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}, as are the undocumented runtime 
exceptions signaled by the Android Platform~\cite{kechagia2014}. 
Some of the findings of our
 study emphasize the impact of these bug hazards on the application robustness 
by mining a different source of information as the ones used in previous works. The present work 
mined issues created by developers on GitHub and Google Code, while  previous research analyzed crash reports and automated test reports.
 Furthermore, our work points to bug hazards that were not detected by previous research (i.e., cross-type wrappings, undocumented checked exceptions and undocumented runtime exceptions thrown by third-party libraries) which represent new threats to application robustness.

@@

The remainder of this paper is organized as follows. 
Section~\ref{sec:back} provides the necessary
background on the Android platform and the Java exception model. 
Section~\ref{sec:study} presents the study design. 
Section~\ref{sec:exceptionminer} describes the ExceptionMiner tool we developed to conduct our study.
Section~\ref{sec:result} reports the findings of our study.
Section~\ref{sec:disc} provides a discussion of the wider implications of our results.
Section~\ref{sec:threats} presents the threats to validity associated to this study. 
Finally Section~\ref{sec:rele} describes related work, 
and Section~\ref{sec:conc} concludes the paper and outlines
directions for future work.
\section{Background}
\label{sec:back}

@@

 In previous work~\cite{coelhoetal15}, we mined exception stack traces available on issues reported on several Android projects hosted in GitHub and Googlecode. To ensure that our sample consists of repositories that were real Android apps, we had to inspect every project site and discard the toy-programs and non-Android repositories. For each repository, we extract the developers whose email are registered on each repository. We emailed the developers and received 68 valid answers.

The majority of our respondents have more than 2 years of Java development experience (87.3\%) and  considerable experience in Android development (more than 1 year)  (85.9\%) - see Figure~\ref{fig:devexpertise}.

\begin{figure} \centering \includegraphics[scale=0.35]{charts/Q1Q2_new.png}
\caption{Developers expertise in Java and Android development.}
\label{fig:devexpertise}
\end{figure}

@@

Moreover, few respondents also answered that such NullPointerException crashes can be caused by backward compatibility as well as memory issues as illustrated in Table X. The other half mentoned that NullPointerExceptions are  common in Android development as it is common in Java standard development.

@@

\bigskip
\noindent\emph{\textbf{RQ 4 Are developers aware of the non-declared checked exceptions?}}
\bigskip

@@

@@

\subsection{Limitations of the Survey-based Study.}

@@

@@

The goal of this paper is to investigate 
to what extent stack trace information can reveal \emph{bug hazards} 
related to exception handling code that may lead to a decrease in
application robustness. 
To that end, we mined the stack 
traces embedded in all issues defined in 482 Android projects hosted in Github and 
157 projects hosted in Google Code. Overall it included 6,005
exception stack traces. Our first key contribution is
a novel approach and toolset (ExceptionMiner) for analyzing Java
  exception stack traces as occurring on GitHub and Google Code
  issues.
Our second contribution is
an empirical study of over 6000 actual stack traces,
  demonstrating that (1) half of the system crashes are due to errors
  in programming logic, with null pointer exceptions being most
  prominent;
  (2) Documentation for explicitly thrown RuntimeExceptions is almost
  never provided; 
  (3) Extensive use of wrapping leads to hard to understand chains
  violating Java's exception handling principles.
Our results shed light on common problems and bug hazards in Java
exception handling code, and call for tool support to help developers
understand their own and third party exception handling and wrapping logic.

\begin{acknowledgements}
If you'd like to thank anyone, place your comments here
and remove the percent signs.
\end{acknowledgements}

>>>>

\title{Exception Handling Bug Hazards in Android}
\subtitle{Results from a Mining and Exploratory Survey Study}

@@

\institute{Roberta Coelho \at
              Federal University of Rio Grande do Norte, CIVT/UFRN. Av. Senador Salgado Filho, 3000. 
Lagoa Nova, CEP: 59.078-970. Natal/RN. Brazil \\
              Tel.: +55-84-3342-2216 \\
              \email{roberta@dimap.ufrn.br}           
           \and
          Lucas Almeida \at
             \email{lucas.almeida@ppgsc.ufrn.br} \\    
          \and    
         Georgios Gousios   \at
  \email{georgios@cs.ru.nl} \\    
 \and
         Arie van Deursen  \at
  \email{arie.vandeursen@tudelft.nl} \\      
\and 
   Christoph Treude \at
  \email{ctreude@dimap.ufrn.br} \\    
}

@@

\begin{abstract}
This paper reports two studies:  a repository mining study whose goal was to identify 
exception handling bug hazards in Android apps; and an exploratory survey study - whose goal
was to assess the perspectives of developers conserning such bug hazards.
The repository mining study mined the exception stack traces embedded on 
159,048 issues reported on Android projects hosted in GitHub (482 projs.) 
and Googlecode (157 projs.).  Overall 6,005 exception stack traces were extracted, 
and analyzed in combination with source code and bytecode analysis. 
Some bug hazards were consistently detected such as:  unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions), and NullPointerExeptions as the major cause of reported chashes in about 50\% of all analyzed projects. Finally, a survey was applied to some developers 
involved on such GitHub projects and their perspective about the bug hazards and 
how they deal with them as assessed and discussed.

\keywords{Exception handling \and Android development \and Repository mining \and Exploratory survey}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

The number of mobile apps is increasing in a daily rate. If on the one hand they
extend phones capabilities far beyond of the basic calls, on the other hand they have to 
cope with an increasing number of exceptinal conditions 
(e.g., faults in underlying middleware or hardware; compatibility issues~\cite{McDon13}; memory and battery restrictions; noisy external resources~\cite{Zhang12}).

Therefore, mechanisms for exception detection and handling are not an optional add-on but a 
fundamental part of such apps. The exception handling mechanism~\cite{goodenough1975exception},
embedded in many mainstream programming languages, such as Java, C++ and C\#,
 is one of the most used techniques for detecting and recovering from such exceptional conditions.
In this paper we will be concerned with exception handling in Android apps,
which reuses Java's exception handling model.

@@

However, such mechanisms are more often then not the less 
understood and tested parts of the system~\cite{miller1997issues,Robil00,shah2010understanding, 
garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling,yuan:2014.osdi}.
As a consequence they may inadvertently negatively affect the system: exception-related code may introduce failures such as 
uncaught exceptions~\cite{jo2004uncaught, Zhang12} - 
which can lead to system crashes, making the system even less robust~\cite{coelho2011unveiling}.

In Java, when an application fails due to an uncaught exception, 
it automatically terminates, while the system prints a stack trace to the console, 
or on a log file~\cite{gosling2000java}.  A typical Java stack trace consists of  the fully qualified name 
of the thrown exception and the ordered list of methods that were active on the call stack before 
the exception occurred~\cite{gosling2000java,bloch2008effective}.
When available, the exception stack traces provide a useful source of information about system crashes ~\cite{bettenburg2008makes} which 
can enable different kinds of post-mortem analysis and support:  debugging~\cite{schroter2010stack}, 
bug classification and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},  
automated bug fixing~\cite{sinha2009fault} and fault-proneness prediction models~\cite{kim2013predicting}.

@@

This work is conducted in two phases. Firstly, a mining study performs a post 
mortem analysis of the exception stack traces 
included in issues reported on Android projects hosted in GitHub and Google Code.
The goal of this study is to investigate whether the reported exception stack traces
 can reveal common \emph{bug hazards} in the exception-related code.  
A \emph{bug hazard}~\cite{binder2000testing}  is a circumstance that increases the 
chance of a bug being present in the software. An example of a bug hazard can 
be a characteristic of  the exception-related code which can increase the likelihood 
of introducing the aforementioned uncaught exceptions.
Secondly, an exploratory survey to the Android developers involved on the mined projects
to assess their perspective about the exception handling bug hazards found.

@@

The exploratory survey was conducted with 70 Android developers involved in one or more of the 
GitHub Android projects whose issues were mined. This survey reveaks that only few developers (3\% of respondents)
 new about the undocumented checked exceptions signaled by native C code of Android platform. 
Moreover, most of the developers recognized that cross-type exception wrappings may
negatively impact the application robustness. The uncaught exceptions due to errors in programming 
logic, with the NullPointerException was identified by most developers (68\%) as the first or second main cause 
of application crashes.

 The high prevalence of NullPointerExceptions found in the mining study and confirmed on the exploratory survey
 is in line with findings of earlier research~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}, 
as are the undocumented runtime exceptions signaled by the Android Platform~\cite{kechagia2014}.

Some of the findings of our mining study emphasize the impact of these bug hazards on the application robustness 
by mining a different source of information as the ones used in previous works. The present work 
mined issues created by developers on GitHub and Google Code, while  previous research analyzed crash reports and automated test reports.
 Furthermore, our work points to bug hazards that were not detected by previous research (i.e., cross-type wrappings, undocumented checked exceptions and undocumented runtime exceptions thrown by third-party libraries) which represent new threats to application robustness.
Moreover, we perform  the first exploratory survey study whose
goal was to assess developers perspective regarding a set of exception handling bug hazards
as well as how developers deal with exceptions and prevent crashes while developing.

@@

The remainder of this paper is organized as follows. 
Section~\ref{sec:back} provides the necessary
background on the Android platform and the Java exception model. 
Section~\ref{sec:study} presents the study design. 
Section~\ref{sec:exceptionminer} describes the ExceptionMiner tool we developed to conduct our study.
Section~\ref{sec:result} reports the findings of our study.
Section~\ref{sec:disc} provides a discussion of the wider implications of our results,
 presents the threats to validity associated to the mining study and discusses the limitations
of the survey-based study. 
Finally Section~\ref{sec:rele} describes related work, 
and Section~\ref{sec:conc} concludes the paper and outlines
directions for future work.
\section{Background}
\label{sec:back}

@@

 In previous work~\cite{coelhoetal15}, we mined exception stack traces available on issues reported on several Android projects hosted in GitHub and Googlecode. To ensure that our sample consists of repositories that were real Android apps, we had to inspect every project site and discard the toy-programs and non-Android repositories. For each repository, we extract the developers whose email are registered on each repository. We emailed the 1,824 developers and received 71 valid answers. The majority of our respondents have more than 2 years of Java development experience (87.3\%) and  considerable experience in Android development (more than 1 year)  (85.9\%) - see Figure~\ref{fig:devexpertise}.

%Q1Q2_new.png
\begin{figure} \centering \includegraphics[scale=0.35]{charts/expertise.png}
\caption{Developers expertise in Java and Android development.}
\label{fig:devexpertise}
\end{figure}

@@

Moreover, few respondents also answered that such NullPointerException crashes can be caused by backward compatibility as well as memory issues as illustrated in Table XXXX. The other half mentoned that NullPointerExceptions are  common in Android development as it is common in Java standard development.

@@

\bigskip
\noindent\emph{\textbf{RQ 4: Are developers aware of the robustness threats caused by JNI undocumented checked exceptions?}}
\bigskip

@@

Developers were asked to rank the main causes of crashes. Approximately 68\% of the repondents ranked the exceptions signaled by programming logic mistakes e.g. NullPointerExceptions as the Top 1 and Top 2 cause of crashes as illustrated in Figure ~\ref{figure:ranking}. In our mining study the exceptions signaled due to programming logic mistakes (java.lang and util) where the causes of 52,0\% of the exception stack traces found. Hence, the developers perspective is aligned with what was revealed in mining study.

\begin{figure*} 
\centering \includegraphics[width=.9\hsize]{charts/maincauses.png}
\caption{Developers perspectives on the main causes of crashes.}\label{fig:ranking} 
\end{figure*}

@@

\subsection{Limitations of the Survey-based Study.}
Due to the exploratory nature of the second phase of this work whose 
goal was to identify the developers perspectives concerning the exception 
handling bug hazards found during the repository mining 
phase, we choose Gounded Theory techniques. The results of our
survey-based study may not apply to every Android developer, since other populations 
might add new insights.  The population we collected data from was comprised 
by Android developers of the GitHub 
Android projects whose issues were mined on the first phase of this work, and who had
time and motivation to answer the survey questions.

Altough the themes and findings that emerged in our study cannot 
be generalized, they give a first view of developers perspectives about
 EH bug hazards found. Hence, we believe that the survey-based study 
have contributed with valuable insights about how Android developers 
used the exception handling code and what are their perspectives regarding
exception handling bug hazards.

@@

\textit{Exploratory Survey Studies.} Exploratory surveys have been used in the software
engineering context to discover the user perspective regarding a broad range of 
topics such as: how developers define productivity [ref]; how GitHub developers use
pull-requests[ref]; how the testing culture of open-shource projects can be characterized [ref]; 
and even how developers use Twiter [ref]. This kind of study is important as a way of 
better understanding the developers behavior and hence provide recommendations and 
tools to help on specific development tasks. This was the first exploratory survey study whose
goal was to assess developers perspective regarding a set of exception handling bug hazards
as well as how developers deal with exceptions and prevent crashes while developing.

@@

The goal of this paper is two-fold: (i) to investigate 
to what extent stack trace information can reveal \emph{bug hazards} 
related to exception handling code that may lead to a decrease in
application robustness; and (ii) investigate the developers perspective
concerning the detected EH \emph{bug hazards}.

To that end, we mined the stack 
traces embedded in all issues defined in 482 Android projects hosted in Github and 
157 projects hosted in Google Code - overall it included 6,005
exception stack traces. And we surveyed developers associated to some of 
the mined GitHub projects.
Our first key contribution is
a novel approach and toolset (ExceptionMiner) for analyzing Java
  exception stack traces as occurring on GitHub and Google Code
  issues.
Our second contribution is
an empirical study of over 6000 actual stack traces,
  demonstrating that (1) half of the system crashes are due to errors
  in programming logic, with null pointer exceptions being most
  prominent;
  (2) Documentation for explicitly thrown RuntimeExceptions is almost
  never provided; 
  (3) Extensive use of wrapping leads to hard to understand chains
  violating Java's exception handling principles.
Our third contribution was the first exploratory to assess the 
developers perspective concerning the exception handling 
bug hazards in Android development.
 Our results shed light on common problems and bug hazards in Java
exception handling code, and call for tool support to help developers
understand their own and third party exception handling and wrapping logic.

%\begin{acknowledgements}
%If you'd like to thank anyone, place your comments here
%and remove the percent signs.
%\end{acknowledgements}
